---
title: "presentation"
output:
  xaringan::moon_reader:
    css: [default, presentation_style.css]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current% / %total%"
    seal: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.retina = 2,
  dpi = 150,
  cache = FALSE
)

library(tidyverse)
library(lubridate)
library(scales)

# ── Palette (Okabe-Ito colorblind-safe, matches output plots) ─────────────────
COL_PSOE    <- "#D55E00"   # vermillion
COL_PP      <- "#0072B2"   # blue
COL_VOX     <- "#009E73"   # bluish green
COL_SUMAR   <- "#CC79A7"   # reddish purple (Okabe-Ito safe)
COL_IMM     <- "#E69F00"   # amber
COL_ACCENT  <- "#332288"   # dark indigo
COL_NEUTRAL <- "#ADB5BD"

SLIDE_THEME <- theme_minimal(base_size = 13) +
  theme(
    plot.background   = element_rect(fill = "transparent", colour = NA),
    panel.background  = element_rect(fill = "white", colour = NA),
    plot.title        = element_text(face = "bold", size = 14,
                                     colour = COL_ACCENT, margin = margin(b = 5)),
    plot.subtitle     = element_text(size = 10, colour = "#555",
                                     margin = margin(b = 8)),
    axis.title        = element_text(size = 10, colour = "#444"),
    axis.text         = element_text(size = 9),
    legend.position   = "bottom",
    legend.title      = element_text(face = "bold", size = 9),
    panel.grid.minor  = element_blank(),
    panel.grid.major  = element_line(colour = "#e8ecf1"),
    plot.margin       = margin(6, 10, 6, 6),
    strip.text        = element_text(face = "bold")
  )
theme_set(SLIDE_THEME)

# ── Load processed data ───────────────────────────────────────────────────────
panel <- read_csv("data/processed/master_panel.csv", show_col_types = FALSE) |>
  mutate(year_month = as.Date(year_month))

polls_raw <- read_csv("data/processed/monthly_polls.csv", show_col_types = FALSE) |>
  mutate(year_month = as.Date(year_month))

forecast_df <- read_csv("data/processed/forecast_2027.csv", show_col_types = FALSE) |>
  mutate(year_month = as.Date(year_month))

model_comp <- read_csv("data/processed/model_comparison.csv", show_col_types = FALSE)

seat_proj <- read_csv("data/processed/seat_projections.csv", show_col_types = FALSE)

seat_mc <- read_csv("data/processed/seat_mc_summary.csv", show_col_types = FALSE)

hist_polls <- tryCatch(
  read_csv("data/processed/historical_polls_2004_2008.csv", show_col_types = FALSE) |>
    mutate(year_month = as.Date(year_month)),
  error = function(e) NULL
)

desc_stats <- read_csv("data/processed/descriptive_stats.csv", show_col_types = FALSE)

# ── New model outputs (match report.Rmd) ──────────────────────────────────────
var_forecast <- tryCatch(
  read_csv("data/processed/var_forecast_2027.csv", show_col_types = FALSE) |>
    mutate(year_month = as.Date(year_month)),
  error = function(e) NULL
)
forecast_accuracy <- tryCatch(
  read_csv("data/processed/forecast_accuracy.csv", show_col_types = FALSE),
  error = function(e) NULL
)
diagnostics_df <- tryCatch(
  read_csv("data/processed/diagnostics_summary.csv", show_col_types = FALSE),
  error = function(e) NULL
)
arima_pred_se <- tryCatch(
  read_csv("data/processed/arima_prediction_se.csv", show_col_types = FALSE),
  error = function(e) NULL
)
pp_psoe_corr_df <- tryCatch(
  read_csv("data/processed/pp_psoe_correlation.csv", show_col_types = FALSE),
  error = function(e) NULL
)

# ── Dynamic variable extraction ───────────────────────────────────────────────
arima_order_label <- if (!is.null(diagnostics_df)) {
  raw <- diagnostics_df |> filter(test == "ARIMA order") |> pull(result) |> first()
  m <- regmatches(raw, regexpr("ARIMA\\([0-9],[0-9],[0-9]\\)", raw))
  if (length(m) == 1) m else "ARIMA(0,1,0)"
} else { "ARIMA(0,1,0)" }

pp_psoe_rho <- if (!is.null(pp_psoe_corr_df)) {
  round(pp_psoe_corr_df$pp_psoe_correlation[1], 2)
} else { 0.72 }

fc_neutral <- forecast_df |> filter(scenario == "Neutral")
fc_neutral_pt <- round(fc_neutral$psoe_forecast[nrow(fc_neutral)], 1)
fc_range_low  <- round(min(forecast_df$psoe_forecast[forecast_df$year_month == max(forecast_df$year_month)]), 1)
fc_range_high <- round(max(forecast_df$psoe_forecast[forecast_df$year_month == max(forecast_df$year_month)]), 1)
```

---
class: title-slide

<div class="title-tag">UC3M &nbsp;·&nbsp; Data Harvesting Course &nbsp;·&nbsp; Spring 2026</div>

# Spanish Government Immigration Rhetoric<br>& Electoral Support

## Multi-Source Data Collection and Statistical Analysis &nbsp;·&nbsp; <span class="title-period">November 2023 – February 2026</span>

<div class="title-stats">
  <div class="title-stat-block"><span class="title-stat-num">4</span><span class="title-stat-label">Sources</span></div>
  <div class="title-divider"></div>
  <div class="title-stat-block"><span class="title-stat-num">28</span><span class="title-stat-label">Months</span></div>
  <div class="title-divider"></div>
  <div class="title-stat-block"><span class="title-stat-num">429</span><span class="title-stat-label">Polls</span></div>
  <div class="title-divider"></div>
  <div class="title-stat-block"><span class="title-stat-num">3,634</span><span class="title-stat-label">Articles</span></div>
  <div class="title-divider"></div>
  <div class="title-stat-block"><span class="title-stat-num">11</span><span class="title-stat-label">Figures</span></div>
  <div class="title-divider"></div>
  <div class="title-stat-block"><span class="title-stat-num">27</span><span class="title-stat-label">Tables</span></div>
</div>

<div class="title-authors">
  <div><span class="title-author-name">Abdullah Tadmuri</span><span class="title-author-id">100502844</span></div>
  <div><span class="title-author-name">Maiheliya Maimaitiming</span><span class="title-author-id">100562134</span></div>
  <div><span class="title-author-name">Tone Varberg Sabri</span><span class="title-author-id">100570811</span></div>
</div>

<div class="title-instructor">Instructor: Patricia Callejo &nbsp;·&nbsp; Universidad Carlos III de Madrid &nbsp;·&nbsp; <code>Rscript main.R</code></div>

---

# Agenda

.pull-left[

**Part I — Data Pipeline**

1. Political context & research question
2. End-to-end pipeline architecture
3. Source 1: La Moncloa — CDX + HTML scraping
4. Source 2: Wikipedia opinion polls
5. Source 3: INE demographics API
6. Source 4: El País media salience
7. 28-month panel overview

]

.pull-right[

**Part II — Key Findings**

8. PSOE support in persistent decline
9. Media salience suppresses support
10. La Moncloa cannot lead the narrative
11. Regression: what actually predicts PSOE?

**Part III — Electoral Projections**

12. `r arima_order_label`-X forecast to June 2027
13. Regularisation law: a non-voter event
14. D'Hondt seat projections (Monte Carlo)
15. Historical parallel: the 2005 Zapatero case

**Conclusions & Limitations**

]

---

# Political Context: Spain 2023–2026

.pull-left[

**November 2023** — Sánchez investiture confirmed (second term), minority government dependent on Junts, ERC, Bildu, PNV

**2024** — Record Canary Islands arrivals via Atlantic route; June EP elections: immigration ranks as top EU voter concern

**2024–2025** — Blocked distribution of unaccompanied minors (*MENA*) between Autonomous Communities triggers inter-regional conflict

**January 2026** — Extraordinary regularisation decree targeting ~500,000 undocumented migrants. First mass legalisation since Zapatero's 2005 process

> **Why this matters:** Spain offers a rare case of a fragile left-wing minority government navigating a politically contested immigration debate under full media scrutiny.

]

.pull-right[

<br>

.stat-row[
<div class="stat-card red"><span class="stat-num">29.1%</span><span class="stat-label">PSOE — Nov 2023</span></div>
<div class="stat-card blue"><span class="stat-num">25.8%</span><span class="stat-label">PSOE — Feb 2026</span></div>
]

<div class="box orange" style="margin-top:1em;">
<strong>Research window:</strong> 28 months of the Sánchez government's second term — from investiture to the regularisation decree
</div>

<div class="box" style="margin-top:0.6em;">
<strong>Key stress-test:</strong> Does communicating actively on immigration help or hurt PSOE? Theory predicts it should help (issue framing). Data will show otherwise.
</div>

]

---

# Research Question & Hypotheses

.pull-left[

> **Does the frequency of official government immigration statements correlate with shifts in PSOE electoral support over the 2023–2026 governing period, and can historical data produce a credible 2027 electoral forecast?**

<br>

| Hypothesis | Direction | Result |
|:-----------|:---------:|:------:|
| H1: Gov. statements → PSOE | Positive | **Not supported** (r = 0.074) |
| H2: Media salience → PSOE | Negative | **Supported** (r = −0.699) |
| H3: Time trend → PSOE | Negative | **Supported** (β = −0.142) |

]

.pull-right[

<br><br>

<div class="box blue" style="margin-bottom:0.6em;">
<strong>H1 (Framing):</strong> Government press releases on immigration reassure centrist voters → PSOE improves
</div>

<div class="box red" style="margin-bottom:0.6em;">
<strong>H2 (Issue ownership):</strong> Higher media salience benefits right-leaning parties via Petrocik (1996) issue ownership mechanism → PSOE falls
</div>

<div class="box orange">
<strong>H3 (Attrition):</strong> Mid-term governmental fatigue is a robust democratic regularity → PSOE falls monotonically
</div>

]

---
class: inverse, center, middle

<div class="section-label">Part I</div>
<div class="section-title">Data Pipeline &<br>Collection Methods</div>
<div class="section-sub">4 sources · 5 scripts · ~10–20 min runtime</div>

---

# End-to-End Pipeline Architecture

<div class="pipeline-grid">

<div class="pipeline-box">
<div class="pipeline-num">Source 01</div>
<div class="pipeline-title">La Moncloa Press Releases</div>
<div class="pipeline-detail">Government immigration statements from Interior, Inclusion & Exteriores ministries<br><br>Nov 2023 – Feb 2026</div>
<span class="pipeline-tech">Wayback CDX API</span>
<span class="pipeline-tech">rvest scraping</span>
<span class="pipeline-tech">httr2</span>
</div>

<div class="pipeline-box b2">
<div class="pipeline-num">Source 02</div>
<div class="pipeline-title">Wikipedia Opinion Polls</div>
<div class="pipeline-detail">Spanish general election polling: PP, PSOE, Vox, Sumar across 429 polls from 47+ firms<br><br>2023–2026</div>
<span class="pipeline-tech">rvest html_table</span>
<span class="pipeline-tech">regex cleaning</span>
<span class="pipeline-tech">monthly agg.</span>
</div>

<div class="pipeline-box b3">
<div class="pipeline-num">Source 03</div>
<div class="pipeline-title">INE Demographics API</div>
<div class="pipeline-detail">Provincial unemployment (EPA, Table 65349) and foreign population trend (Table 9683)<br><br>Quarterly</div>
<span class="pipeline-tech">INE JSON API</span>
<span class="pipeline-tech">httr2 / jsonlite</span>
<span class="pipeline-tech">47 provinces</span>
</div>

<div class="pipeline-box b4">
<div class="pipeline-num">Source 04</div>
<div class="pipeline-title">El País Media Salience</div>
<div class="pipeline-detail">Immigration topic-section headlines from El País; normalised 0–100 salience index<br><br>3,634 articles</div>
<span class="pipeline-tech">Paginated scraping</span>
<span class="pipeline-tech">polite delays</span>
<span class="pipeline-tech">keyword filter</span>
</div>

</div>

<br>

<div class="box" style="font-size:0.85em;">
<strong>Integration:</strong> All four sources are aggregated to a <strong>28-month monthly panel</strong> (2023-11 to 2026-02) in <code>data/processed/master_panel.csv</code> — the unit of analysis for all models.
</div>

---

# Source 1: La Moncloa — CDX API + HTML Scraping

.pull-left[

**The problem:** La Moncloa's index page uses JavaScript-rendered filtering — URL parameters are ignored by the server, returning today's content regardless.

**The solution:** Wayback Machine CDX API to discover historical press release URLs, then live scraping to extract body text.

**Two-stage architecture:**
1. CDX API discovers all archived URLs matching immigration slug keywords from three ministries
2. Live scraping of confirmed URLs extracts title, body text, word count, and date
3. Body text matched against 25 immigration keywords to confirm relevance

**Output:** 64 confirmed immigration press releases over 28 months (mean: 2.3/month, max: 8 in a single month)

.muted[robots.txt checked via `robotstxt` package; polite 2–4 s delays applied]

]

.pull-right[

```r
# Stage 1 — CDX API URL discovery
CDX_BASE <- "https://web.archive.org/cdx/search/cdx"

cdx_url <- paste0(
  CDX_BASE,
  "?url=www.lamoncloa.gob.es/ministerios/",
  ministry, "/*",
  "&output=json",
  "&limit=5000",
  "&from=", START_DATE,   # "20231101"
  "&to=",   END_DATE,     # "20260301"
  "&filter=statuscode:200"
)

cdx_resp <- request(cdx_url) |>
  req_headers(`User-Agent` = USER_AGENT) |>
  req_retry(max_tries = 3, backoff = ~ 8) |>
  req_perform()

urls <- resp_body_json(cdx_resp) |>
  map_chr(~ .x[[3]])  # column 3 = original URL

# Stage 2 — Live HTML scraping (XPath)
page <- read_html(url)

# XPath: find <h1> anywhere in the DOM tree
title <- page |>
  xml_find_first("//h1") |>
  xml_text(trim = TRUE)

# XPath union: try multiple body containers
body_nodes <- page |>
  xml_find_all(paste(
    "//div[contains(@class,'noticia-texto')]//p",
    "//div[contains(@class,'cuerpo-noticia')]//p",
    "//article//p",
    sep = " | "))
body <- xml_text(body_nodes) |>
  paste(collapse = " ")

# Regex: extract date from URL slug
date <- str_extract(slug, "^\\d{6,8}")
```

]

---

# Source 2: Wikipedia Opinion Polls

.pull-left[

**Target:** Wikipedia article *"Opinion polling for the next Spanish general election"*

**Page structure:** 5 wikitables (one per year + historical CIS series), sorted newest-first, with polling firm, dates, sample size, and vote shares per party

**Challenges:**
- Header rows use image `alt` attributes, not plain text → extracted via `html_attr("alt")`
- Percentage columns contain ranges (e.g., `"28.1–29.3%"`) → midpoint extraction via regex
- Fieldwork dates span two rows → cleaned with `lubridate::dmy()`
- 47+ polling firms with inconsistent name formatting → standardised

**Output:** 429 polls → aggregated to 28 monthly averages

]

.pull-right[

```r
WIKI_URL <- paste0(
  "https://en.wikipedia.org/wiki/",
  "Opinion_polling_for_the_next_",
  "Spanish_general_election"
)

page <- read_html(
  request(WIKI_URL) |>
    req_headers(`User-Agent` = UA) |>
    req_retry(max_tries = 3,
              backoff = ~ 5) |>
    req_perform() |>
    resp_body_string()
)

# Extract all wikitables
tables <- page |>
  html_elements("table.wikitable")

# XPath: party names hidden in <img alt>
header_ths <- tables[[1]] |>
  xml_find_all(".//tr[1]/th")
col_names <- map_chr(header_ths, \(th) {
  txt <- xml_text(th, trim = TRUE)
  if (nchar(txt) > 0) return(txt)
  # Fallback: image alt attribute
  th |> xml_find_first(".//img") |>
    xml_attr("alt") %|% "unknown"
})

# Regex: extract % from range strings
pct <- str_extract(raw_polls$PSOE,
  "\\d+\\.?\\d*") |> as.numeric()
```

]

---

# Source 3: INE Demographics API

.pull-left[

**The INE (Instituto Nacional de Estadística)** provides a REST/JSON API with no authentication required.

**Two series collected:**

**1. Unemployment by province** — EPA Table 65349 (Operation 293), quarterly, 477 series covering all 47 Spanish provinces by sex. Used to construct provincial unemployment context for the reaction map.

**2. Foreign population trend** — Table 9683, filtered to total non-Spanish population at national level (sex: ambos sexos). Provides the structural demographic backdrop to the immigration debate.

**Rate limiting:** INE API is tolerant; `req_retry(max_tries = 3, backoff = ~ 10)` used as a precaution. No `robots.txt` restrictions.

.muted[Documented at: ine.es/dyngs/DAB — publicly available open data]

]

.pull-right[

```r
INE_BASE <- "https://servicios.ine.es/wstempus/js/ES"

# Helper: fetch any INE table as JSON
fetch_ine <- function(endpoint, params = "") {
  url <- paste0(INE_BASE, "/", endpoint,
    if (nchar(params) > 0)
      paste0("?", params) else "")

  resp <- request(url) |>
    req_headers(`User-Agent` = UA) |>
    req_retry(max_tries = 3,
              backoff = ~ 10) |>
    req_timeout(30) |>
    req_perform()

  resp |> resp_body_json()
}

# Foreign population — Table 9683
# tv=70032:16420  → total nationality
# tv=70013:451    → ambos sexos (both)
pop_data <- fetch_ine(
  "DATOS_TABLA/9683",
  "tv=70032:16420,tv=70013:451"
)

# Unemployment — EPA Table 65349
unemp <- fetch_ine("DATOS_TABLA/65349")
```

]

---

# Source 4: El País Media Salience

.pull-left[

**Target:** El País immigration section — paginated topic pages covering all tagged immigration articles

**Approach:** Paginated scraping of `elpais.com/tag/inmigracion/` with polite delays; headlines and dates extracted from article cards

**Salience construction:**
1. Raw monthly article count (mean: 128 articles/month, range: 51–208)
2. Keyword filter: 27 immigration-related terms applied to headline text
3. Normalised 0–100 index: `n_immigration / max(n_immigration) × 100`

**Result:** Salience index captures media agenda intensity, not just volume — months with 208 articles score 100; months with 51 score 24.5

.muted[El País blocks headless browsers; standard `httr2` GET with browser User-Agent works within robots.txt rules]

]

.pull-right[

```{r salience-bar, fig.width=4.8, fig.height=3.0, out.width="100%"}
salience <- read_csv("data/processed/salience_scores.csv",
                     show_col_types = FALSE) |>
  mutate(year_month = as.Date(year_month),
         imm_volume_norm = round(
           n_immigration / max(n_immigration, na.rm = TRUE) * 100, 1))

ggplot(salience, aes(year_month, imm_volume_norm)) +
  geom_col(aes(fill = imm_volume_norm > 75),
           width = 25, show.legend = FALSE) +
  scale_fill_manual(values = c("#c9d8e8", COL_ACCENT)) +
  scale_x_date(date_labels = "%b\n%Y", date_breaks = "4 months") +
  scale_y_continuous(labels = function(x) paste0(x)) +
  labs(title = "El País Immigration Salience Index (0–100)",
       x = NULL, y = "Salience score") +
  theme(plot.title = element_text(size = 11))
```

]

---

# 28-Month Panel Dataset

.pull-left[

```{r desc-table}
desc_stats |>
  mutate(across(where(is.numeric), ~ round(., 2))) |>
  knitr::kable(
    col.names = c("Variable", "N", "Mean", "SD", "Min", "Median", "Max"),
    format = "html"
  ) |>
  kableExtra::kable_styling(font_size = 12, full_width = TRUE) |>
  kableExtra::row_spec(0, background = "#1D3557", color = "white", bold = TRUE) |>
  kableExtra::row_spec(c(1, 2), background = "#eef3f9")
```

]

.pull-right[

<br>

<div class="stat-row">
<div class="stat-card"><span class="stat-num">28</span><span class="stat-label">Monthly obs.</span></div>
<div class="stat-card blue"><span class="stat-num">429</span><span class="stat-label">Polls scraped</span></div>
</div>

<div class="stat-row" style="margin-top:0.4em;">
<div class="stat-card orange"><span class="stat-num">64</span><span class="stat-label">Press releases</span></div>
<div class="stat-card red"><span class="stat-num">3,634</span><span class="stat-label">Articles</span></div>
</div>

<div class="box" style="margin-top:0.9em; font-size:0.88em;">
Panel spans <strong>2023-11-01 to 2026-02-01</strong>. All four sources merged by <code>year_month</code> (first of month). No months missing.
</div>

<div class="box orange" style="font-size:0.88em;">
PSOE polling firms: 47+ agencies including CIS, GAD3, Metroscopia, Key Data, Sigma Dos. Monthly averages weighted by recency within the polling window.
</div>

]

---
class: inverse, center, middle

<div class="section-label">Part II</div>
<div class="section-title">Key Findings</div>
<div class="section-sub">Three results that define the relationship between<br>immigration communication and electoral support</div>

---

# Finding 1: PSOE Support in Persistent Decline

```{r timeline, fig.width=10.0, fig.height=2.8, out.width="100%"}
panel_long <- panel |>
  select(year_month, PSOE = psoe_avg, PP = pp_avg, Vox = vox_avg) |>
  pivot_longer(-year_month, names_to = "Party", values_to = "support")

ggplot(panel_long, aes(year_month, support, colour = Party)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 1.8) +
  geom_smooth(data = filter(panel_long, Party == "PSOE"),
              method = "lm", se = TRUE, linetype = "dashed",
              colour = COL_PSOE, fill = COL_PSOE, alpha = 0.12, linewidth = 0.7) +
  annotate("rect",
           xmin = as.Date("2026-01-01"), xmax = as.Date("2026-02-28"),
           ymin = -Inf, ymax = Inf,
           fill = COL_IMM, alpha = 0.12) +
  annotate("text", x = as.Date("2026-01-15"), y = 34.5,
           label = "Jan 2026\nRegularisation", size = 2.9,
           colour = "#9b6000", hjust = 0.5, lineheight = 0.9) +
  scale_colour_manual(values = c(PP = COL_PP, PSOE = COL_PSOE, Vox = COL_VOX)) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
  scale_y_continuous(labels = function(x) paste0(x, "%"), limits = c(8, 38)) +
  labs(title = "Voting Intention: PSOE, PP and Vox — Nov 2023 to Feb 2026",
       subtitle = "Monthly averages from 429 polls · Dashed line = OLS trend for PSOE",
       x = NULL, y = "Vote intention (%)", colour = NULL) +
  theme(legend.position = "right",
        legend.text = element_text(size = 9))
```

.pull-left[
<div class="box red" style="font-size:0.85em; margin-top:0.4em;">
<strong>PSOE trend:</strong> −0.142 pp/month (OLS, p &lt; 0.001) · Cumulative observed fall: <strong>≈ 3.3 pp</strong> over 28 months
</div>
]

.pull-right[
<div class="box blue" style="font-size:0.85em; margin-top:0.4em;">
<strong>PP lead:</strong> PP averaged 30.3% vs PSOE 27.2% · PP-PSOE gap widened from ~5 pp to ~8 pp by end of period
</div>
]

---

# Finding 2: Media Salience Suppresses PSOE Support

.pull-left[

```{r scatter, fig.width=4.8, fig.height=3.2, out.width="100%"}
panel_col <- panel |>
  mutate(period = case_when(
    year_month < as.Date("2024-07-01") ~ "Early (Nov 23–Jun 24)",
    year_month < as.Date("2025-07-01") ~ "Mid  (Jul 24–Jun 25)",
    TRUE                               ~ "Late (Jul 25–Feb 26)"
  ),
  period = factor(period, levels = c("Early (Nov 23–Jun 24)",
                                     "Mid  (Jul 24–Jun 25)",
                                     "Late (Jul 25–Feb 26)")))

ggplot(panel_col, aes(imm_volume_norm, psoe_avg)) +
  geom_point(aes(colour = period), size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = TRUE,
              colour = COL_ACCENT, fill = COL_ACCENT, alpha = 0.1,
              linewidth = 0.9) +
  scale_colour_manual(values = c("#56B4E9", COL_ACCENT, COL_PSOE)) +
  scale_x_continuous(labels = function(x) paste0(x)) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(title = "El País Salience vs PSOE Support",
       subtitle = paste0("Each point = one month   ·   r = ", round(cor(panel_col$imm_volume_norm, panel_col$psoe_avg, use = "complete.obs"), 3)),
       x = "Immigration salience (0–100 norm.)",
       y = "PSOE vote intention (%)",
       colour = NULL) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8))
```

]

.pull-right[

<div class="box red" style="margin-bottom:0.6em; margin-top:0.3em;">
<strong>Pearson r = −0.699</strong> (p &lt; 0.001)<br>
Months with highest immigration media coverage consistently coincide with lowest PSOE polling.
</div>

<div class="box orange" style="margin-bottom:0.7em;">
<strong>OLS coefficient:</strong> β = −0.014 (NW-robust p ≈ 0.10) — marginal significance after controlling for time trend. Effect is real but partially mediated by the secular decline.
</div>

<div class="box blue">
<strong>Mechanism (Petrocik 1996):</strong> Issue ownership theory — when immigration becomes more salient, voters increasingly associate the issue with right-leaning parties (PP, Vox), not the governing left.
</div>

]

---

# Finding 2b: The Correlation Is Structural, Not Episodic

.pull-left[

```{r rolling-corr, fig.width=4.8, fig.height=3.2, out.width="100%"}
roll_data <- panel |>
  arrange(year_month) |>
  filter(!is.na(psoe_avg), !is.na(imm_volume_norm))

roll_data$roll_corr <- zoo::rollapplyr(
  cbind(roll_data$psoe_avg, roll_data$imm_volume_norm),
  width = 6, FUN = function(mat) cor(mat[,1], mat[,2], use = "complete.obs"),
  by.column = FALSE, fill = NA_real_
)

# 95% significance threshold for Pearson r with n = 6 (df = 4)
df_roll <- 4
t_crit  <- qt(0.975, df = df_roll)
ci_roll <- t_crit / sqrt(t_crit^2 + df_roll)

ggplot(roll_data |> filter(!is.na(roll_corr)),
       aes(x = year_month, y = roll_corr)) +
  geom_ribbon(aes(ymin = -ci_roll, ymax = ci_roll),
              fill = COL_NEUTRAL, alpha = 0.25) +
  geom_hline(yintercept = 0, colour = COL_NEUTRAL, linewidth = 0.5) +
  geom_line(colour = COL_PSOE, linewidth = 1.1) +
  geom_point(colour = COL_PSOE, size = 2, shape = 21,
             fill = "white", stroke = 1.1) +
  annotate("text", x = max(roll_data$year_month, na.rm = TRUE),
           y = -ci_roll - 0.08, label = "p = 0.05",
           colour = "grey50", size = 2.5, hjust = 1) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %y") +
  scale_y_continuous(limits = c(-1.05, 1.05), breaks = seq(-1, 1, 0.5)) +
  labs(title = "Rolling Correlation: PSOE vs Immigration Salience",
       subtitle = sprintf("6-month window \u00b7 Shaded = 95%% significance (|r| > %.2f)", ci_roll),
       x = NULL, y = "Pearson r") +
  theme(plot.title = element_text(size = 11))
```

]

.pull-right[

<br>

<div class="box red" style="margin-bottom:0.7em;">
<strong> Highly unstable.</strong> The rolling correlation is initially  positive, however it  reverses sharply negative by Autumn 2024, recovers again to +0.75 in early 2025, before declining into persistent negative territory from mid-2025 onward.
</div>

<div class="box orange" style="margin-bottom:0.7em;">
<strong>Deepening trend:</strong> The correlation is negative toward Nov 2025, and stabilizes from Nov 2025&ndash;Feb 2026 (regularisation debate), suggesting the relationship strengthens as the issue intensifies.
</div>

<div class="box blue">
<strong>Implication:</strong> The observation window is relatively short, which limits our ability to distinguish structural long-term effects from common temporal trends. As a result, the observed association may reflect a persistent relationship, but it could also be driven by broader time dynamics that affect both variables during the study period
</div>

]

---

# Finding 3: La Moncloa Cannot Lead the Narrative

.pull-left[

**Granger causality tests** (order = 2, Newey-West robust) test whether past values of X help predict future PSOE support beyond what PSOE's own history predicts.

| Predictor | F-stat | p-value | Verdict |
|:----------|:------:|:-------:|:-------:|
| Gov. statements | 0.138 | 0.872 | **No predictive power** |
| Media salience | 0.342 | 0.714 | **No predictive power** |

<br>

**Cross-correlation peak:** No lag exceeds the significance threshold (±0.377). The largest CCF value is at lag 3 (r = −0.297), still well below the bound. Bivariate Pearson r between statements and PSOE levels = 0.074 (near-zero).

<div class="box" style="margin-top:0.7em; font-size:0.88em;">
<strong>Implication:</strong> La Moncloa's communication strategy is <em>reactive</em>, not strategic — press release volume responds to media salience, not the reverse.
</div>

]

.pull-right[

<br>

<div class="box red" style="margin-bottom:0.8em;">
<strong>Correlation: r = 0.074</strong><br>
Government immigration statements are essentially uncorrelated with PSOE voting intention. The framing hypothesis (H1) is comprehensively rejected.
</div>

<div class="box orange" style="margin-bottom:0.8em;">
<strong>Reactive vs strategic:</strong> Press release count shows weak positive correlation with media salience (r = 0.101) — ministries may respond to the news cycle, but the association is negligible.
</div>

<div class="box blue">
<strong>Policy implication:</strong> Increasing the frequency of immigration press releases is unlikely to arrest the polling decline. The dominant predictor is time — not communication.
</div>

]

---

# Regression: What Actually Predicts PSOE Support?

.pull-left[

```{r model-compare, fig.width=4.8, fig.height=2.8, out.width="100%"}
mc <- model_comp |>
  mutate(
    short_name = case_when(
      str_detect(model, "M1") ~ "M1: Trend only",
      str_detect(model, "M2") ~ "M2: + Gov. statements",
      str_detect(model, "M3") ~ "M3: Full model (main)",
      str_detect(model, "M4") ~ "M4: Lagged",
      TRUE ~ model
    ),
    short_name = factor(short_name, levels = rev(c(
      "M1: Trend only", "M2: + Gov. statements",
      "M3: Full model (main)", "M4: Lagged"
    ))),
    best = str_detect(model, "M3")
  )

ggplot(mc, aes(r.squared, short_name, fill = best)) +
  geom_col(width = 0.52, show.legend = FALSE) +
  geom_text(aes(label = sprintf("R\u00b2 = %.3f", r.squared)),
            hjust = -0.1, size = 3.2, fontface = "bold",
            colour = COL_ACCENT) +
  scale_fill_manual(values = c("grey78", COL_ACCENT)) +
  scale_x_continuous(limits = c(0, 0.96),
                     labels = function(x) sprintf("%.2f", x),
                     expand = expansion(mult = c(0, 0))) +
  coord_cartesian(clip = "off") +
  labs(title = "Model Comparison by R\u00b2",
       x = "R-squared", y = NULL) +
  theme(axis.text.y  = element_text(size = 10, face = "bold"),
        plot.margin  = margin(6, 55, 6, 6))
```

]

.pull-right[

| | M1 | M3 ★ |
|:--|:--:|:--:|
| R² | 0.695 | **0.727** |
| Adj. R² | 0.684 | **0.693** |
| AIC | 77.6 | 78.5 |
| n | 28 | 28 |

<div class="box" style="margin-top:0.6em; font-size:0.84em;">
<strong>M3 coefficients (NW-robust):</strong><br>
• Time trend: β = −0.142, p &lt; 0.001 ★★★<br>
• Gov. statements: β = −0.055, p = 0.674 (n.s.)<br>
• Media salience: β = −0.014, p ≈ 0.10 ·
</div>

<div class="box orange" style="font-size:0.84em; margin-top:0.5em;">
M1→M3: R² improves by only <strong>3.2 pp</strong>. AIC barely changes (77.6→78.5). The dominant driver is <strong>time trend</strong> — immigration variables add negligible explanatory power.
</div>

]

---
class: inverse, center, middle

<div class="section-label">Part III</div>
<div class="section-title">Electoral Projections</div>
<div class="section-sub">Forecast · Regularisation analysis · Seat projections · Historical parallel</div>

---

# `r arima_order_label`-X Forecast: PSOE to June 2027

```{r forecast-plot, fig.width=10.0, fig.height=4.0, out.width="100%"}
hist_panel <- panel |>
  select(year_month, psoe_avg) |>
  mutate(type = "Observed", scenario = "Observed")

scen_colors <- c(
  "Neutral"                         = COL_ACCENT,
  "High salience (regularization debate)" = COL_PSOE,
  "Low salience (issue fades)"       = COL_VOX
)

fcst_plot <- forecast_df |>
  filter(year_month >= as.Date("2026-03-01"))

ggplot() +
  geom_line(data = hist_panel, aes(year_month, psoe_avg),
            colour = "#555", linewidth = 1.05) +
  geom_point(data = hist_panel, aes(year_month, psoe_avg),
             colour = "#555", size = 1.5) +
  geom_ribbon(data = fcst_plot,
              aes(year_month, ymin = lower_95, ymax = upper_95,
                  fill = scenario), alpha = 0.10) +
  geom_ribbon(data = fcst_plot,
              aes(year_month, ymin = lower_80, ymax = upper_80,
                  fill = scenario), alpha = 0.18) +
  geom_line(data = fcst_plot,
            aes(year_month, psoe_forecast, colour = scenario),
            linewidth = 1.0) +
  geom_vline(xintercept = as.Date("2026-03-01"),
             linetype = "dashed", colour = COL_IMM, linewidth = 0.8) +
  annotate("text", x = as.Date("2026-04-01"), y = 29.5,
           label = "Forecast \u2192", size = 3, colour = COL_IMM, hjust = 0) +
  # Direct endpoint labels on the last forecast point (Jun 2027)
  geom_text(data = fcst_plot |> filter(year_month == max(year_month)),
            aes(x = year_month, y = psoe_forecast, colour = scenario,
                label = paste0(round(psoe_forecast, 1), "%")),
            hjust = -0.15, size = 3, fontface = "bold", show.legend = FALSE) +
  scale_colour_manual(values = scen_colors) +
  scale_fill_manual(values = scen_colors) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "4 months",
               expand = expansion(mult = c(0.02, 0.10))) +
  scale_y_continuous(labels = function(x) paste0(x, "%"), limits = c(18, 33)) +
  labs(title = paste0("PSOE Voting Intention: Observed + ", arima_order_label, "-X Forecast by Scenario"),
       subtitle = "Shaded bands = 80% and 95% prediction intervals   \u00b7   Forecast horizon: Mar 2026 \u2013 Jun 2027",
       x = NULL, y = "PSOE vote intention (%)",
       colour = "Scenario", fill = "Scenario") +
  theme(legend.position = "right", legend.text = element_text(size = 8))
```

.pull-left[
<div class="box red" style="font-size:0.84em; margin-top:0.3em;">
<strong>June 2027:</strong> Neutral ≈ `r fc_neutral_pt`% · High salience ≈ `r fc_range_low`% · Low salience ≈ `r fc_range_high`%
</div>
]
.pull-right[
<div class="box orange" style="font-size:0.84em; margin-top:0.3em;">
Spread across scenarios: <strong>&lt; 0.5 pp</strong> — the dominant driver is the time trend, not the salience scenario
</div>
]

---

# The January 2026 Regularisation: A Non-Voter Event

.pull-left[

**The regularisation decree** grants legal status to ~500,000 undocumented migrants who meet residency criteria. Media framing portrays this as electorally significant.

**The data says otherwise:**

| Pathway | Adds voters? | Why? |
|:--------|:------------:|:----:|
| Legal residency only | ❌ | Residency ≠ citizenship |
| Citizenship via naturalization | ❌ | Requires 10 yrs (general) / 2 yrs (Ibero-American) |
| 2027 General Election suffrage | ❌ | Only Spanish nationals may vote |
| Referendum / EU Parliament | ❌ | EU residence rights require citizenship |

<div class="box red" style="font-size:0.88em; margin-top:0.7em;">
<strong>Zero direct electoral impact.</strong> No regularised migrant will gain Spanish suffrage by the June 2027 General Election under any realistic timeline.
</div>

]

.pull-right[

<br>

<div class="box orange" style="margin-bottom:0.7em;">
<strong>Where the effect IS:</strong> Issue salience. The regularisation debate sustained high El País immigration coverage through Jan–Feb 2026 — the highest salience months in the study window.
</div>

<div class="box blue" style="margin-bottom:0.7em;">
<strong>Indirect electoral pathways:</strong><br>
• Mobilises existing left-wing voters who support the measure<br>
• Triggers right-wing counter-mobilisation (PP + Vox)<br>
• Keeps immigration high on the agenda → favours issue-ownership holders (PP/Vox per H2)
</div>

<div class="box" style="font-size:0.85em;">
<strong>Historical parallel (2005):</strong> Zapatero's regularisation (~700,000 applications) coincided with a sustained PSOE recovery — but that recovery was driven by macro-economic conditions, not the regularisation per se.
</div>

]

---

# D'Hondt Seat Projections — Monte Carlo (n = 5,000)

.pull-left[

```{r seats-plot, fig.width=4.8, fig.height=3.0, out.width="100%"}
seat_long <- seat_proj |>
  pivot_longer(-c(scenario, Total), names_to = "Party", values_to = "Seats") |>
  mutate(
    scenario = factor(scenario,
      levels = c("High salience (regularization debate)",
                 "Neutral",
                 "Low salience (issue fades)")),
    Party = factor(Party, levels = c("PP", "PSOE", "Vox", "Sumar", "Regional")),
    pty_col = case_when(
      Party == "PP"       ~ COL_PP,
      Party == "PSOE"     ~ COL_PSOE,
      Party == "Vox"      ~ COL_VOX,
      Party == "Sumar"    ~ COL_SUMAR,
      Party == "Regional" ~ COL_NEUTRAL
    )
  )

ggplot(seat_long, aes(Seats, scenario, fill = Party)) +
  geom_col(position = "dodge", width = 0.65) +
  geom_vline(xintercept = 176, linetype = "dashed",
             colour = "black", linewidth = 0.7, alpha = 0.6) +
  annotate("text", x = 177, y = 0.5, label = "176\nMajority",
           size = 2.8, hjust = 0, vjust = 0, lineheight = 0.85) +
  scale_fill_manual(values = c(
    PP = COL_PP, PSOE = COL_PSOE, Vox = COL_VOX,
    Sumar = COL_SUMAR, Regional = COL_NEUTRAL
  )) +
  scale_x_continuous(limits = c(0, 200)) +
  scale_y_discrete(labels = c(
    "High salience (regularization debate)" = "High salience",
    "Neutral" = "Neutral",
    "Low salience (issue fades)" = "Low salience"
  )) +
  labs(title = "D'Hondt Seat Projections by Scenario",
       subtitle = "Point estimates from June 2027 vote share forecast",
       x = "Seats (350 total)", y = NULL, fill = NULL) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8))
```

]

.pull-right[

<br>

| Party | High | Neutral | Low |
|:------|:----:|:-------:|:---:|
| **PP** | **145** | **144** | **142** |
| PSOE | 113 | 114 | 115 |
| Vox | 34 | 34 | 34 |
| Sumar | 26 | 26 | 27 |
| Regional | 32 | 32 | 32 |

<br>

<div class="box red" style="font-size:0.86em; margin-bottom:0.6em;">
<strong>PP absolute majority probability (Monte Carlo):</strong><br>
High salience: <strong>71.7%</strong> &nbsp;|&nbsp; Neutral: <strong>46.1%</strong> &nbsp;|&nbsp; Low: <strong>21.3%</strong>
</div>

<div class="box orange" style="font-size:0.86em;">
<strong>Coalition arithmetic (Neutral scenario):</strong><br>
PP + Vox = 144 + 34 = **178** (majority)<br>
PSOE + Sumar + Regional = 114 + 26 + 32 = **172** (short)<br>
PP alone falls short of 176 — Vox cooperation is decisive.
</div>

]

---

# Historical Parallel: The 2005 Zapatero Regularisation

.pull-left[

```{r historical, fig.width=4.8, fig.height=3.2, out.width="100%"}
if (!is.null(hist_polls)) {
  hist_polls <- hist_polls |>
    mutate(months_rel = as.numeric(
      difftime(year_month, as.Date("2005-02-01"), units = "weeks") / 4.33
    ))

  # 2026 overlay: current PSOE data aligned to Jan 2026 regularisation
  current_overlay <- panel |>
    filter(!is.na(psoe_avg)) |>
    mutate(months_rel = as.numeric(
      difftime(year_month, as.Date("2026-01-01"), units = "weeks") / 4.33
    ))

  ggplot() +
    # 2005 counterfactual ribbon
    geom_ribbon(data = hist_polls,
                aes(x = months_rel, ymin = psoe_counterfact, ymax = psoe_pct),
                fill = COL_PP, alpha = 0.10) +
    geom_line(data = hist_polls, aes(months_rel, psoe_counterfact),
              linetype = "dashed", colour = "#aaa", linewidth = 0.7) +
    # 2005 actual
    geom_line(data = hist_polls, aes(months_rel, psoe_pct),
              colour = COL_PSOE, linewidth = 1.0) +
    geom_point(data = hist_polls, aes(months_rel, psoe_pct),
               colour = COL_PSOE, size = 1.5) +
    # 2026 overlay
    geom_line(data = current_overlay, aes(months_rel, psoe_avg),
              colour = COL_ACCENT, linewidth = 1.0, linetype = "solid") +
    geom_point(data = current_overlay, aes(months_rel, psoe_avg),
               colour = COL_ACCENT, size = 1.5, shape = 17) +
    # Regularisation marker
    geom_vline(xintercept = 0, linetype = "dashed",
               colour = COL_IMM, linewidth = 0.9) +
    annotate("text", x = 0.5, y = max(hist_polls$psoe_pct, na.rm = TRUE) - 0.5,
             label = "Regularisation\nstart", size = 2.5,
             hjust = 0, colour = "#9b6000", lineheight = 0.9) +
    # Legend labels
    annotate("text", x = max(hist_polls$months_rel, na.rm = TRUE) - 1,
             y = min(hist_polls$psoe_pct, na.rm = TRUE) + 1.5,
             label = "2005 (Zapatero)", colour = COL_PSOE,
             size = 2.8, hjust = 1, fontface = "bold") +
    annotate("text", x = max(current_overlay$months_rel, na.rm = TRUE),
             y = min(current_overlay$psoe_avg, na.rm = TRUE) - 1,
             label = "2026 (S\u00e1nchez)", colour = COL_ACCENT,
             size = 2.8, hjust = 1, fontface = "bold") +
    scale_y_continuous(labels = function(x) paste0(x, "%")) +
    scale_x_continuous(breaks = seq(-25, 25, 5),
                       labels = function(x) paste0(ifelse(x >= 0, "+", ""), x, "m")) +
    labs(title = "PSOE Support: 2005 vs 2026 Regularisation Comparison",
         subtitle = "Aligned on regularisation start month (t = 0)",
         x = "Months relative to regularisation",
         y = "PSOE vote intention (%)") +
    theme(plot.title = element_text(size = 11))
} else {
  ggplot() +
    annotate("text", x = 0.5, y = 0.5, label = "Historical data not available",
             size = 5, colour = "#aaa") +
    theme_void()
}
```

]

.pull-right[

<br>

**ITS (Interrupted Time Series) results:**

| Term | Estimate | NW p-value |
|:-----|:--------:|:----------:|
| Baseline (Apr 2004) | 42.4% | &lt; 0.001 |
| Pre-trend (t) | −0.486 pp/m | &lt; 0.001 |
| Immediate shift | −1.13 pp | 0.101 (n.s.) |
| Post-trend change | **+0.493 pp/m** | &lt; 0.001 |

<div class="box green" style="font-size:0.86em; margin-top:0.7em;">
After the 2005 regularisation, PSOE's declining trend <strong>reversed</strong>: the party recovered at +0.493 pp/month, winning the 2008 election at <strong>43.9%</strong>.
</div>

<div class="box orange" style="font-size:0.86em;">
<strong>2026 caveat:</strong> Zapatero governed in a strong economic environment (GDP growth >3%). Sánchez faces fragmented parliament + elevated cost-of-living concerns. Direct comparison is limited.
</div>

]

---

# Course Techniques Applied

| Technique | Where Used | Example |
|:----------|:-----------|:--------|
| **HTML table scraping** | Script 02 (Wikipedia) | `html_table(fill = TRUE)` on 5 wikitables |
| **XPath expressions** | Scripts 01, 02, 04 | `xml_find_first("//h1")`, `xml_find_all("//div[contains(@class,'noticia-texto')]//p \| ...//p")` |
| **CSS selectors** | Scripts 01–04 | `html_elements("table.wikitable")`, `html_element("time")` |
| **Regex (stringr)** | All scripts | `str_extract(slug, "^\\d{6,8}")`, `str_detect(text, "inmigra")` |
| **REST/JSON APIs** | Script 03 (INE) | `httr2::request()` → `resp_body_json()` → `jsonlite` parsing |
| **Paginated scraping** | Scripts 01, 04 | Loop over CDX pages and El País `/noticias/inmigracion/N/` |
| **Ethical scraping** | All scripts | `robotstxt` check, `Sys.sleep(3–5)`, standard User-Agent |

<div class="box green" style="margin-top:0.6em; font-size:0.86em;">
<strong>All seven core Data Harvesting topics</strong> appear at least once. XPath and CSS selectors are used side-by-side across all four data sources.
</div>

---
class: inverse, center, middle

<div class="section-label">Part IV</div>
<div class="section-title">Conclusions</div>
<div class="section-sub">Five findings with direct evidence</div>

---

# Five Key Conclusions

<div class="finding"><div class="finding-num">1</div><div><strong>PSOE support fell −0.142 pp/month throughout the Sánchez term</strong> — a cumulative observed decline of ~3.3 pp from 29.1% (Nov 2023) to ~25.8% (Feb 2026). The trend is robust across all model specifications (adj. R² = 0.695–0.705).</div></div>

<div class="finding"><div class="finding-num">2</div><div><strong> Higher media attention to immigration is associated with lower support for the Spanish Socialist Workers' Party (r = −0.699, p < 0.001). Consistent with the issue-ownership theory of John R. Petrocik, greater salience may shift the debate toward an issue where parties like the People's Party and Vox are seen as stronger.
However, this pattern may partly reflect a shared time trend. After removing the trend, the effect becomes weaker and only marginally significant (p ≈ 0.10).</div></div>

<div class="finding"><div class="finding-num">3</div><div><strong>La Moncloa's immigration communication has zero predictive power</strong>. Granger tests (F = 0.138, p = 0.872 for statements; F = 0.342, p = 0.714 for salience) confirm no lead-lag effect. The government reacts to the news cycle; it does not shape it.</div></div>

<div class="finding"><div class="finding-num">4</div><div><strong>The January 2026 regularisation adds zero direct voters</strong>. All ~500,000 beneficiaries remain ineligible for national suffrage through 2027. Electoral consequences are entirely mediated via issue salience — which our model shows hurts PSOE.</div></div>

<div class="finding"><div class="finding-num">5</div><div><strong>2027 projections place PP around 136–137 seats and PSOE around 120 seats. Immigration salience shifts the margin slightly (≈3 seats) but does not materially change coalition outcomes.</div></div>


---

# Limitations & Reproducibility

.pull-left[

**Technical limitations**

- La Moncloa CDX coverage: Wayback Machine indexes ~85–90% of pages; some press releases may be missing
- Wikipedia polls: CIS polling methodology changed in 2024 — pre/post comparison requires caution
- El País paywall: only headlines and metadata accessible without subscription — article body sentiment not captured
- INE provincial unemployment: quarterly frequency → interpolated to monthly (linear)

**Modelling limitations**

- n = 28 monthly observations — low statistical power for multi-predictor OLS
- `r arima_order_label` assumes a random walk; model selection via AIC/BIC on short series is inherently uncertain
- Omitted variables: unemployment shocks, coalition crises, EU elections not modelled

]

.pull-right[

**Reproducibility**

All data collection and analysis is fully reproducible:

```r
# Full pipeline: ~10–20 minutes
Rscript main.R

# Render report
Rscript -e "rmarkdown::render('report.Rmd')"
```

Outputs:
- `data/raw/` — all scraped raw files
- `data/processed/` — cleaned panel + model results
- `output/plots/` — all figures
- `report.html` — full self-contained report

<div class="box green" style="margin-top:0.8em; font-size:0.86em;">
<strong>Pipeline verified:</strong> All 8 scripts run end-to-end with no manual steps. Data refresh requires only re-running <code>main.R</code>.
</div>

]

.footnote[UC3M · Data Harvesting Course · Spring 2026 · Abdullah Tadmuri, Maiheliya Maimaitiming, Tone Varberg Sabri]
